Compressive Channel Sensing of Multipath Millimeter Channel via a Beamspace Dantzig Selector and its Probabilitic Performance Bound

### 2018-12-31 ongoing/report/article.tex
Outline theorem proofs

In this article, we utilize a hybrid structure at both transmitter and receiver end, consisting of both digital and analog percoders and combiners.
We argue in favor of DS, and verify that DS is, as originally perceived, ideal in certain senses.
DS encompasses a more general setting than OMP, such as in its requirement of restricted isometry property (RIP) only, and as in its allowance of noise corruption.
Moreover, we extend their proof to derive a stronger bound in the setting of estimation of millimeter wave channel for almost-sparse channel matrix

@@

Let the direct product
%
\begin{gather}
\Omega
=\Omega_t \times \Omega_h \times \Omega_n \times \Omega_r
\end{gather}
%
denote the overall sample space, where:
%
\begin{itemize}
\item \(\Omega_t\) denotes the collection of probability samples at the transmitter end,
\item \(\Omega_h\) denotes that of the communication channel,
\item \(\Omega_h\) denotes that of the noise arising in the communication channel,
\item \(\Omega_r\) denotes that at the receiver end.
\end{itemize}
%
\(\omega_1\) stands for the probability index of the transmitter end, and \(\omega_2\) for the channel, and \(\omega_3\) for the receiver.

@@

@@

And consider the virtual representation of the MIMO channel (see for example Akdeniz et.\ al.\ 2014).
\begin{gather}
H
=\sum_{l=0}^L \alpha_l \mathbf{a}(\phi_l) \mathbf{a}(\theta_l)^\dagger
\end{gather}
The physical meaning of \(\theta_l\) is the \(l\)-th angle of arrival, and \(\theta_l\) the \(l\)-th angle of arrival.
It is noted that in real scenario, the exponent \(i n \phi_l\) should be multiplied by a factor \((1/\lambda) \sin \eta\), \(\eta\) being the incident angle, accounting for the spacing of two adjacent antenna.
But, since we consider uniform distribution of \(\theta_l, \phi_l\)'s, it suffices to absorb such factors into the values of \(\theta_l, \phi_l\)'s.

If we absorb angle of incidence (formed by the ray of electronic wave and the normal line)
\((d/\lambda) \sin \tilde{\theta}_l\)
Moreover, let the convention for discrete Fourier transform be adopted as below for completeness of presentation.

\myCount
\textbf{Definition.} Let \(\mathbf{v} \in \mathbb{R}^m\) be given.
We say \(u \in \mathbb{R}^m\) is the discrete Fourier transform of \(\mathbf{v}\), if
\begin{gather}
\mathbf{u} =F \mathbf{v}
\end{gather}
where each entry of the matrix \(F\) is defined to be
\begin{gather}
F_{m,n} =e^{2\pi ikm/N},
\quad m,n =0,1,2,\dotsc, N-1
\end{gather}

In summary, the effective channel may be written as
\begin{gather}
Y (\omega_h, \omega_z)
:=W_B W_R ( H(\omega_h) F_R F_B \mathbf{u}_m +z(\omega_z) )
\end{gather}
Supress the probabilistic index for simplicity, we restate the same equation as
\begin{gather}
Y
:=W_B W_R ( H F_R F_B \mathbf{u}_m +z )
\end{gather}

@@

The effective channel \(Y\) has \(N\) columns, but to estimate it, we may agree to send one pilot signal at one time, which consists of only one antenna sending signal, namely \(\mathbf{u}_n\).
This way, we may denote the acquired effective channel matrix \(Y\) under noisy observation as (borrowing Matlab's colon notation to denote submatrix)
\begin{gather}
Y_{:,n}
:=W_B W_R ( H F_R F_B \mathbf{u}_m +z ), \quad
n =1, 2, \dotsc, N_B
\end{gather}
The crux of the matter is the recovery of \(H\) with knowledge of \(Y\), and of course \(W_R, W_B, F_R, F_B\) which are in our control.

as sparse recovery problem of \(\mathrm{vec}(H)\).
But that suffers the problem of grid quantization, and 
On the other hand, the author argues in favor of DS, since it is provably ideal in certain senses, guaranteed even under noisy condition,
So why do we not simply formulate the recovery of \(H\) as a DS program?
DS is an \(\ell_1\)-minimization problem, and, since it is well known to be convex, it is no less tractable than OMP.

In addition, observe that, if we write
\begin{gather}
G =K^\dagger H
\end{gather}
as the spatial frequency domain representation of \(H\), then
\begin{gather}
Y
:=W_B W_R K ( G F_R F_B +K^\dagger z ), \quad
n =1, 2, \dotsc, N_B
\end{gather}
If at the end of day, \(G\) is sparse, then we recover \(G\) instead, and our use of DS is fully justified.

That is to say, \(K^\dagger z\) may be absorbed, and a similar Kronecker product may be used.
\begin{gather}
\mathrm{vec}(Y)
\approx \big[ (K F_R^\dagger F_B^\dagger) \otimes (W_B W_R) \big] \mathrm{vec}(H)
\end{gather}
If so, our program reads

\myCount
\textbf{Program}.
Let \(H(\omega_h, \omega_n)\) be the virtual channel, and \(W_R(\omega_r)\) the random analog combiner, \(W_B(\omega_r)\) the random analog combiner, \(F_R(\omega_t)\) the random analog combiner, \(F_B(\omega_t)\) the random analog combiner, and \(K\) the discrete Fourier matrix, all as described in previous sections.
Suppose \(Y\) is the acquired effective channel matrix under noisy observation.
Set the beamspace vector representation of \(H\) to be
\begin{gather}
h := \mathrm{vec}(H) \\
\tilde{y} := \mathrm{vec}(Y) \\
A =(K^\dagger F_R F_B) \otimes (W_B W_R K)
\end{gather}
and set
\begin{gather}
r =\tilde{y} -A h
\end{gather}
Then, with parameter \(\gamma\) specified, define the following program
\begin{gather}
\hat{\tilde{g}}
=\underset {{h}} {\mathrm{min}} \|h\|_1 \quad
\mathrm{s.t.}\;\; \|A^\dagger r\|_\infty \leq \gamma
\end{gather}
And convert \(\hat{\tilde{g}}\) back to the space domain as
\begin{gather}
\hat{h}
=K^\dagger \hat{\tilde{g}}.
\end{gather}
And finally recover the estimated \(H\) as
\begin{gather}
\hat{H} =\mathrm{vec}^{-1} (\hat{h})
\end{gather}

It will not be hard to establish Cand\`es and Tao's result to the setting of millimeter wave virtual channel, and resulting bound is immediate under straighforward work, as below.

@@

\myCount
\textbf{Lemma.} Let \(\phi(\omega_3)\) be a uniformly distributed in \([0,2\pi)\) according to probability space index \(\omega_3 \in \Omega\).
And the function \(\mathbf{a}(\phi)\) is defined as in \eqref{}.
Then the discrete Fourier transform

To find the \(l\)-th path's contribution to the frequency domain, recall the identity
\begin{align}
S(\phi)
:=&\left| \sum_{n=0}^{N-1} e^{i n \phi} \right| \\ \notag
=&\left| \frac{e^{i N \phi} -1}{e^{i \phi} -1} \right| \\ \notag
=&\left| \frac{e^{i N \phi/2} -e^{-i N \phi/2}}{e^{i \phi /2} -e^{-i \phi /2}} \cdot e^{i (N-1) \phi/2} \right| \\ \notag
=&\frac{|\sin(N \phi/2)|}{|\sin(\phi /2)|},
\quad 0 \leq \phi \leq \pi
\end{align}
By taking the first two term of the expansion \(\sin x =x -x^3/6 +- \dotsc\), and noting the monotony, 
\begin{align}
S(\phi)
\leq& \frac{1}{\phi/2 -(\phi/2)^3/6} \\ \notag
=& \frac{-48}{\phi^3 -24\phi}
\quad 0 \leq \phi \leq \pi,
\end{align}

@@

By the same token, more generally,
\begin{align}
&\left| \frac{1}{N} \sum_{n=0}^{N-1} e^{-2\pi i k n/N} e^{i n \phi} \right| \\ \notag
=&\frac{|\sin( \phi N /2 )|}{N |\sin( \pi k /N -\phi /2 )|} \\ \notag
=&\frac{1}{N} S(\pi \phi') \\ \notag
\leq& \frac{48}{N \pi^3 (\phi'^3 -24 \phi' /\pi^2)}
\end{align}
where \(\phi' =k/N -\phi/2\pi\), and the cubic expression is split.
We wish to bound the tail of \(S(\pi \phi')\), and apply the na\"ive bound
\begin{align}
&\frac{N}{2\pi} \cdot \frac{48}{N \pi^3} \int_{L/N}^1 \frac{1}{\phi'^3 -24 \phi' /\pi^2}
 d\phi' \\ \notag
=&\frac{1}{2\pi^2} \cdot \left( 2 \log \frac{L}{N} +\log \frac{24-\pi^2}{24-\pi^2 L^2 /N^2} \right) \\ \notag
\leq& \frac{1}{\pi^2} \log \frac{L}{N}
\end{align}
Here, the integral in the first line is immediate.
The second term in the bracket is smaller than 0 and is dropped in the last inequality.
The bound is satisfying enough, though refinement is surely possible.
Indeed, \(\log L/N \to 0\) as \(N \to \infty\) and limited by the number of paths \(L\).

%A =\mathrm{supp} \beta

where
\begin{gather}
\mathcal{A}, \mathcal{B}, \mathcal{C} \subset \mathbb{N} \\ \notag
\mathcal{A} \cap \mathcal{B} =B \cap C =A \cap C =\varnothing \\ \notag
A \cup B \cup C =\{1, \dotsc, N\}.
\end{gather}

\textbf{Proposition}.
Let \(h \in \mathbb{R}^n\) and \(\eta \in \mathbb{R}^p\) are defined as in \eqref{}.
Then
\begin{align}
\|h_A\|_1 -\|\eta_A\|_1 +\|\eta_C\|_1 -\|h_C\|_1
\leq& \|h +\eta\|_1 \\ \notag
\leq& \|h\|_1
\end{align}

\textbf{Proof}.

\begin{align}
\|X^\dagger_{AB} X \eta\|_\infty
<&\|X^\dagger X \eta\|_\infty \\ \notag
\leq& 2 \lambda_p
\end{align}

From ``Dantzig Selector'' (Cand\`es and Tao 2007), Lemma 1, first equation.

\Result
{Lemma}
{
\begin{align}
\|\eta_Ah\|_2
\leq K_1 \|X^\dagger_{AB} X \eta\|_2 +K_2 \|\eta_C\|_2
\end{align}
}

From ``Dantzig Selector'' (Cand\`es and Tao 2007), Lemma 1, second equation.

\Result
{Lemma}
{
\begin{align}
\|\eta\|_2^2
\leq \|\eta_{AB}\|_2^2 +K_3 \|\eta_C\|_2^2
\end{align}
}

>>>>

In this article, we utilize a hybrid structure at both transmitter and receiver end, each of them having both digital and analog percoders and combiners.
We argue in favor of DS, and verify that DS is, as originally perceived, ideal in certain senses.
DS encompasses a more general setting than OMP, such as in its requirement of restricted isometry property (RIP) only, and as in its allowance of noise corruption.
Moreover, we extend their proof to derive a stronger bound in the setting of estimation of millimeter wave channel for almost-sparse channel matrix

@@

@@

Since expression involving probability will get very complecated, introduce the direct product
%
\begin{gather}
\Omega
=\Omega_t \times \Omega_h \times \Omega_n \times \Omega_r
\end{gather}
%
to denote the overall sample space, where
%
\begin{itemize}
\item \(\Omega_t\) denotes the collection of probability samples at the transmitter end,
\item \(\Omega_h\) denotes that of the communication channel,
\item \(\Omega_h\) denotes that of the noise arising in the communication channel,
\item \(\Omega_r\) denotes that at the receiver end.
\end{itemize}
%
\(\omega_1\) stands for the event space index of the transmitter end, and \(\omega_2\) for the channel, and \(\omega_3\) for the receiver.

@@

And consider the virtual representation of the MIMO channel (see for example Akdeniz et.\ al.\ 2014).
\begin{gather}
H
=\sum_{l=0}^L \alpha_l \mathbf{a}\big( \frac{d}{\lambda} \sin \phi_l' \big) \mathbf{a}\big( \frac{d}{\lambda} \sin \theta_l' \big)^\dagger
\end{gather}
The physical meaning of \(\theta_l\) is the \(l\)-th angle of incidence (formed by the ray and the normal line) of departure electronic wave, and \(\theta_l\), the \(l\)-th that of arrival wave.

For our purpose, we may absorb the argument of \(\mathbf{a}\) as
\begin{gather}
\phi_l
=\frac{d}{\lambda} \sin \phi_l' \\
\theta_l
=\frac{d}{\lambda} \sin \theta_l'
\end{gather}
to get a simpler form
\begin{gather}
H
=\sum_{l=0}^L \alpha_l \mathbf{a}(\phi_l) \mathbf{a}(\theta_l)^\dagger
\end{gather}

Moreover, let the convention for discrete Fourier transform be adopted as below for completeness of presentation.

\Result
{Definition.}
{
Let \(\mathbf{v} \in \mathbb{R}^m\) be given.
We say \(u \in \mathbb{R}^m\) is the discrete Fourier transform of \(\mathbf{v}\), if
\begin{gather}
\mathbf{u} =\dagger F \mathbf{v}
\end{gather}
where each entry of the matrix \(K\) is defined to be
\begin{gather}
   K(m,n) =\frac{1}{\sqrt{N}} e^{2\pi imn/N},
\quad m,n =0,1,2,\dotsc, N-1
\end{gather}
where \(i =\sqrt{-1}\).
}

\subsection{System Model}

We consider the situation with a hybrid structure at both transmitter and receiver end.
Each end consists of both digital and analog percoders and combiners.
In the transmitter end, there are digital precoder \(F_B\) and analog precoder \(F_R\).
Similarly, in the receiver end, there are digital combiner \(W_B\) and analog combiner \(W_R\).
Recall that analog precoder is restricted to have value with magnitude being unity.
To set up notation, we list them as below.
\begin{gather}
F_B \in \mathbb{R}^{N_r \cdot N_s} \\
F_R \in \mathbb{R}^{N_h \cdot N_r}, |F_R(i,j)| =1 \\
F_B \in \mathbb{R}^{N_r \cdot N_h} \\
F_B \in \mathbb{R}^{N_s \cdot N_r}, |F_R(i,j)| =1.
\end{gather}
And we consider zero-mean, unity-variance additive white Gaussian noise \(z\).
Since we restrict our discussion to a specific slice of time, the signal \(z\) degrades as a entrywise i.i.d.\ standard normal distribution vector.

In summary, the effective channel may be written as
\begin{gather}
U (\omega_h, \omega_e)
:=W_B W_R ( H(\omega_h) F_R F_B \mathbf{u}_m +z(\omega_z) )
\end{gather}
Supress the probabilistic index for simplicity, we restate the same equation as
\begin{gather}
Y
:=W_B W_R ( H F_R F_B \mathbf{u}_m +z )
\end{gather}

@@

The effective channel \(U\) has \(N\) columns, but to estimate it, we may agree to send one pilot signal at one time, which consists of only one antenna sending signal, namely \(\mathbf{u}_n\).
This way, we may denote the acquired effective channel matrix \(U\) under noisy observation as (borrowing Matlab's colon notation to denote submatrix)
\begin{gather}
Y_{:,n}
:=W_B W_R ( H F_R F_B \mathbf{u}_m +z ), \quad
n =1, 2, \dotsc, N_B
\end{gather}
The crux of the matter is the recovery of \(H\) with knowledge of \(U\), and of course \(W_R, W_B, F_R, F_B\) which are in our control.

Following previous work on CSS, we write
\begin{gather}
h := \mathrm{vec}(H) \\
u := \mathrm{vec}(U) \\
A =(K^\dagger F_R F_B) \otimes (W_B W_R K)
\end{gather}
to formulate our goal as sparse recovery problem of \(\mathrm{vec}(H)\).
But, as we have motivated in previous sections, we want to elimitate the angle grid quantization problem, and we want to use DS by considering sparse vector.
What shall we do?

Observe that, if we write
\begin{gather}
G =K^\dagger H
\end{gather}
as the spatial frequency domain representation of \(H\), then
\begin{gather}
Y
:=W_B W_R K ( G F_R F_B +K^\dagger z ), \quad
n =1, 2, \dotsc, N_B
\end{gather}
If at the end of day, \(G\) is sparse, then we recover \(G\) instead, and our use of DS is fully justified.
To the best of our knowledge, no work has explicitly considered the beamspace CSS.

That is to say, \(K^\dagger z\) may be absorbed, and a similar Kronecker product may be used.
\begin{gather}
\mathrm{vec}(Y)
\approx \big[ (F_B^\dagger F_R^\dagger) \otimes (W_B W_R K) \big] \mathrm{vec}(H)
\end{gather}
To simplify notation, set 
\begin{gather}
P =(F_B^\dagger F_R^\dagger) \otimes (W_B W_R K)
\end{gather}
The instances of \(P\) is completely determined in the transmitter and receiver end, that is
\begin{gather}
P =P(\omega_t, \omega_r)
\end{gather}
If so, our program reads

\Result
{Program}
{
Let \(H(\omega_h, \omega_n)\) be the virtual channel, and \(W_R(\omega_r)\) the random analog combiner, \(W_B(\omega_r)\) the random analog combiner, \(F_R(\omega_t)\) the random analog combiner, \(F_B(\omega_t)\) the random analog combiner, and \(K\) the discrete Fourier matrix, all as described in previous sections.
Suppose \(U\) is the acquired effective channel matrix under noisy observation.
\(A\) is defined just above in ().
Set
\begin{gather}
r =u -P h
\end{gather}
Then, with parameter \(\gamma\) specified, define the following program
\begin{gather}
\hat{g}
=\underset {{h}} {\mathrm{min}} \|h\|_1 \quad
\mathrm{s.t.}\;\; \|A^\dagger r\|_\infty \leq \gamma
\end{gather}
And convert \(\hat{g}\) back to the space domain as
\begin{gather}
\hat{h}
=K^\dagger \hat{g}.
\end{gather}
And finally recover the estimated \(H\) as
\begin{gather}
\hat{H} =\mathrm{vec}^{-1} (\hat{h})
\end{gather}
}

It will not be hard to establish Cand\`es and Tao's result to the setting of millimeter wave virtual channel, and resulting bound is immediate under straighforward work, as follows.

@@

We follow the argument outlined in ``The Dantzig Selector'' and fill the content.
The sparse vector to be recovered of our investigation is \(g \in \mathbb{R}^{N_H^2}\), and we continue to use \(\hat{g}\) to denote the Dantzig Selector.

We first establish a technical lemma that shows \(\mathrm{supp}(g)\) has roughly cardinal \(L\).
At the heart of the argument, the lemma provides the basis on which \(g\) is almost-sparse.

Consider
\begin{gather}
d =g -\hat{g}
\end{gather}
And let
\begin{itemize}
\item \(\mathcal{A}\) be the largest \(L\) position of \(g\).
\item \(\mathcal{B}\) be the largest \(L\) position of \(d\) except on \(\mathcal{A}\).
\item \(\mathcal{C} =\{1, \dotsc, N_H^2\} -( \mathcal{B} \cup \mathcal{C} )\), i.e., the complementary index set.
\end{itemize}
So \(\mathcal{A}, \mathcal{B}, \mathcal{C}\) are mutually exclusive index sets by construction:
\begin{gather}
\mathcal{A}, \mathcal{B}, \mathcal{C} \subset \{1, \dotsc, N_H^2\} \\
\mathcal{A} \cap \mathcal{B} =B \cap C =A \cap C =\varnothing \\
A \cup B \cup C =\{1, \dotsc, N\}.
\end{gather}

\Result
{Lemma.}
{
Let \(\phi(\omega_h)\) and \(\theta(\omega_h)\) be both uniformly, independently distributed in \([0,2\pi)\), and beamspace channel matrix \(G\) defined as in ().
\begin{gather}
\|g\|_1 \leq \frac{1}{\pi^2} \log \frac{L}{N}.
\end{gather}

To find the \(l\)-th path's contribution to the frequency domain, recall the identity for Dirichlet kernel
\begin{align}
S(\phi)
:=&\left| \sum_{n=0}^{N-1} e^{i n \phi} \right| \\ \notag
=&\frac{|\sin(N \phi/2)|}{|\sin(\phi /2)|},
\quad 0 \leq \phi \leq \pi
\end{align}
By taking the first two term of the expansion \(\sin x =x -x^3/6 +- \dotsc\), and noting the monotony, 
\begin{align}
S(\phi)
\leq& \frac{1}{\phi/2 -(\phi/2)^3/6} \\ \notag
=& \frac{-48}{\phi^3 -24\phi}
\quad 0 \leq \phi \leq \pi,
\end{align}

@@

By the same token, more generally,
\begin{align}
&\left| \frac{1}{N} \sum_{n=0}^{N-1} e^{-2\pi i k n/N} e^{i n \phi} \right| \\ \notag
=&\frac{|\sin( \phi N /2 )|}{N |\sin( \pi k /N -\phi /2 )|} \\ \notag
=&\frac{1}{N} S(\pi \phi') \\ \notag
\leq& \frac{48}{N \pi^3 (\phi'^3 -24 \phi' /\pi^2)}
\end{align}
where \(\phi' =k/N -\phi/2\pi\), and the cubic expression is split.
We wish to bound the tail of \(S(\pi \phi')\), and apply the na\"ive bound
\begin{align}
&\frac{N}{2\pi} \cdot \frac{48}{N \pi^3} \int_{L/N}^1 \frac{1}{\phi'^3 -24 \phi' /\pi^2}
 d\phi' \\ \notag
=&\frac{1}{2\pi^2} \cdot \left( 2 \log \frac{L}{N} +\log \frac{24-\pi^2}{24-\pi^2 L^2 /N^2} \right) \\ \notag
\leq& \frac{1}{\pi^2} \log \frac{L}{N}
\end{align}
Here, the integral in the first line is immediate.
The second term in the bracket is smaller than 0 and is dropped in the last inequality.
}

The bound is satisfying enough, though refinement is surely possible.
Indeed, \(\log L/N \to 0\) as \(N \to \infty\) and limited by the number of paths \(L\).

The rest follows ``The Dantzig Selector'' very closely.

\Result
{Proposition}
{
Let \(h \in \mathbb{R}^n\) and \(\eta \in \mathbb{R}^p\) are defined as in ().
Then
\begin{align}
\|d_{BC}\|_1
\leq \|d_A\|_1 +\|g_{BC}\|_1
\end{align}
}
To show this, observe
\begin{align}
\|g\|_1 -\|d_A\|_1 +\|d_{BC}\|_1 -\|g_{BC}\|_1
\leq& \|g +d\|_1 \\ \notag
\leq& \|g\|_1
\end{align}
In the first line, we recall \(g =g_A\).

\Result
{Lemma}
{
\begin{align}
|\langle e, P_{:,j} \rangle| \leq \sqrt{2 \log p}
\end{align}
}

\Result
{Lemma}
{
\begin{align}
\|P^\dagger P d\|_\infty
\leq& 2 \sqrt{2 \log p}.
\end{align}
}
To show this, start with
\begin{align}
\langle e -r, P_{:,j} \rangle
=\langle P \hat{g} -P g, P_{:,j} \rangle
=\langle P g, P_{:,j} \rangle
\end{align}
Result follows from triangle inequality, and Lemma ().

From ``Dantzig Selector'' (Cand\`es and Tao 2007), Lemma 1, first equation, we have the result below.

\Result
{Lemma}
{
\begin{align}
\|d_A\|_2
\leq \frac{1}{1-\delta} \|P^\dagger_{AB} P \eta\|_2 +\frac{\theta}{(1-\delta) \sqrt{L}} \|d_{BC}\|_1
\end{align}
}

From ``Dantzig Selector'' (Cand\`es and Tao 2007), Lemma 1, second equation, we have the result below.

\Result
{Lemma}
{
\begin{align}
\|d\|_2^2
\leq \|d_{AB}\|_2^2 +\frac{1}{L} \|d_C\|_1^2
\end{align}
}

We can formulate our main theorem:

(To be done)
